---
title: "Human_Fragmentomics"
output: html_document
date: "2024-05-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Packages

```{r}

library(data.table)
library(dplyr)
library(faraway)
library(ggExtra)
library(GGally)
library(ggsci)
library(ggplot2)
library(ggpubr)
library(ggthemr)
library(gridExtra)
library(gt)
library(gtsummary)
library(rlang)
library(scales)
library(survival)
library(survminer)
library(readxl)
library(reshape2)
library(tidyverse)

ggthemr("fresh")


load("/Users/graceliu/bratman/data_frames_compressed.RData")
```

# Data processing

```{r eval=FALSE,include=FALSE,echo=FALSE}

df_OPC <- read_excel("/Users/lucaspenny/Documents/1 Bratman/head_neck_analysis_lucas/summary/Quantification/summ_OPC.v7.1.xlsx")
colnames(df_OPC)[2] = c("patient")

df_clinical = fread("/Users/lucaspenny/Documents/1 Bratman/head_neck_analysis_lucas/clincalData/clinical_data.tsv")

df_gtv_nodes = fread("~/Documents/1 Bratman/head_neck_analysis_lucas/clincalData/df_gtv_nodes.tsv")

df_gtv_primary = fread("~/Documents/1 Bratman/head_neck_analysis_lucas/clincalData/df_gtv_primary.tsv")

df_OPC = df_OPC %>% 
  left_join(df_clinical,by=("patient")) %>% 
  left_join(select(df_gtv_nodes,patient,sum_gtv_nodes), by = "patient") %>% 
  left_join(select(df_gtv_primary,patient,gtv_primary), by = "patient")

df_OPC$Sample <- gsub("_L0\\d+.*", "", df_OPC$Sample)

df_hg_healthy_frag = fread("~/Documents/1 Bratman/head_neck_analysis_lucas/HPV-seq/data/hg/fragmentLength/20240130_HUCON_fragmenthLength.csv") 
df_hg_healthy_frag$Sample <- gsub("_insert_size_metrics.txt", "", 
                                  df_hg_healthy_frag$filename) 
df_hg_healthy_frag = df_hg_healthy_frag[1:20,]

df_jinfeng_fraglength = fread("/Users/lucaspenny/Documents/1 Bratman/head_neck_analysis_lucas/HPV-seq/data/hg/fragmentLength/20240513_jinfeng_hg_fraglength.csv")
df_jinfeng_fraglength$Sample <- gsub("_L0\\d+.*", "", df_jinfeng_fraglength$filename)

df_prop_jinfeng <- df_jinfeng_fraglength %>%
  # Calculate the sum of columns 3 through 603
  mutate(total_sum = rowSums(.[, 53:253])) %>%
  dplyr::select(Sample, everything()) %>%
  mutate(across(53:253, ~ . / total_sum)) %>% 
  dplyr::select(Sample,53:253)  
  # filter(Sample == "OPC_Pool-5_62_S62") %>% 
  # pivot_longer(cols = 2:202, names_to = "Variable", values_to = "Value")

# fwrite(df_prop_jinfeng,"~/Desktop/human_fragmentlength_proportions.txt",quote=F,row.names = F,sep="\t",col.names = T)


# df_hg_frag = fread("~/Documents/1 Bratman/head_neck_analysis_lucas/HPV-seq/data/hg/fragmentLength/20240513_lucas_hg_fragment_length_summary.csv")
# colnames(df_hg_frag)[2] = c("Sample")

# df_prop_lucas <- df_hg_frag %>%
#   # Calculate the sum of columns 3 through 603
#   mutate(total_sum = rowSums(.[, 53:253])) %>%
#   dplyr::select(Sample, everything()) %>%
#   mutate(across(53:253, ~ . / total_sum)) %>% 
#   dplyr::select(Sample,53:253) 
# 
df_prop_healthy <- df_hg_healthy_frag %>%
  # Calculate the sum of columns 3 through 603
  mutate(total_sum = rowSums(.[, 53:253])) %>%
  dplyr::select(Sample, everything()) %>%
  mutate(across(53:253, ~ . / total_sum)) %>%
  dplyr::select(Sample,53:253)



save(df_clinical, df_gtv_nodes, df_gtv_primary, df_hg_healthy_frag, df_jinfeng_fraglength, df_OPC, df_prop_healthy, df_prop_jinfeng, file = "/Users/lucaspenny/Documents/1 Bratman/head_neck_analysis_lucas/HPV-seq/data/20240624_NatMed_dfs/data_frames_compressed.RData", compress = TRUE)

# attach('/Users/lucaspenny/Documents/1 Bratman/head_neck_analysis_lucas/HPV-seq/data/HPV_analysis_dataframes.Rdata'); df_gtv_nodes <- df_ctdna_eqd2_radiomic_clinical
# df_ctdna_eqd2_radiomic_clinical <- df_ctdna_eqd2_radiomic_clinical
```

The basic workflow i'm going to use is select the sample ID's that relate to the specific condition and then make the ggplot from that.

# Basic Visualization of features

## T- Category

```{r ,fig.width=15,fig.height=4}
#################
# N Category = N0 
#################
stat_tcat = df_OPC %>% 
  filter(stage_t_ajcc8 == "T1" & Timepoint == "Baseline") 

df_test = df_prop_jinfeng %>% 
  filter(Sample %in% stat_tcat$Sample) %>% 
  pivot_longer(cols = 2:202, names_to = "Variable", values_to = "Value") %>% 
  mutate(Variable = as.integer(Variable)) 

# Calculating means, standard deviations, and min/max values
df_summary = df_test %>%
  group_by(Variable) %>%
  summarize(
    Mean = mean(Value),
    SD = sd(Value),
    SE = SD / sqrt(n()),  # Standard Error
    Min = min(Value),
    Max = max(Value)
  )

# ggplot(df_test,aes(x=Variable,y=Value,group=Sample)) +
#   geom_line() + labs(
#     x = "Fragment Length (bp)",
#     y = "Proportion")


a = ggplot(df_test, aes(x=Variable, y=Value, group=Sample)) +
  geom_line(alpha=0.4) + 
  geom_line(data = df_summary, aes(x=Variable, y=Mean, group=1), size=1, color="black") + 
  labs(x = "Fragment Length (bp)", y = "Proportion")

#################
# N Category = N1 
#################
stat_tcat = df_OPC %>% 
  filter(stage_t_ajcc8 == "T2" & Timepoint == "Baseline") 

df_test = df_prop_jinfeng %>% 
  filter(Sample %in% stat_tcat$Sample) %>% 
  pivot_longer(cols = 2:202, names_to = "Variable", values_to = "Value") %>% 
  mutate(Variable = as.integer(Variable)) 

# Calculating means, standard deviations, and min/max values
df_summary = df_test %>%
  group_by(Variable) %>%
  summarize(
    Mean = mean(Value),
    SD = sd(Value),
    SE = SD / sqrt(n()),  # Standard Error
    Min = min(Value),
    Max = max(Value)
  )

# ggplot(df_test,aes(x=Variable,y=Value,group=Sample)) +
#   geom_line() + labs(
#     x = "Fragment Length (bp)",
#     y = "Proportion")


b = ggplot(df_test, aes(x=Variable, y=Value, group=Sample)) +
  geom_line(alpha=0.4) + 
  geom_line(data = df_summary, aes(x=Variable, y=Mean, group=1), size=1, color="black") + 
  labs(x = "Fragment Length (bp)", y = "Proportion")

#################
# N Category = N1 
#################
stat_tcat = df_OPC %>% 
  filter(stage_t_ajcc8 == "T3" & Timepoint == "Baseline") 

df_test = df_prop_jinfeng %>% 
  filter(Sample %in% stat_tcat$Sample) %>% 
  pivot_longer(cols = 2:202, names_to = "Variable", values_to = "Value") %>% 
  mutate(Variable = as.integer(Variable)) 

# Calculating means, standard deviations, and min/max values
df_summary = df_test %>%
  group_by(Variable) %>%
  summarize(
    Mean = mean(Value),
    SD = sd(Value),
    SE = SD / sqrt(n()),  # Standard Error
    Min = min(Value),
    Max = max(Value)
  )

# ggplot(df_test,aes(x=Variable,y=Value,group=Sample)) +
#   geom_line() + labs(
#     x = "Fragment Length (bp)",
#     y = "Proportion")


c = ggplot(df_test, aes(x=Variable, y=Value, group=Sample)) +
  geom_line(alpha=0.4) + 
  geom_line(data = df_summary, aes(x=Variable, y=Mean, group=1), size=1, color="black") + 
  labs(x = "Fragment Length (bp)", y = "Proportion")


#################
# N Category = N3 
#################
stat_tcat = df_OPC %>% 
  filter(stage_t_ajcc8 == "T4" & Timepoint == "Baseline") 

df_test = df_prop_jinfeng %>% 
  filter(Sample %in% stat_tcat$Sample) %>% 
  pivot_longer(cols = 2:202, names_to = "Variable", values_to = "Value") %>% 
  mutate(Variable = as.integer(Variable)) 

# Calculating means, standard deviations, and min/max values
df_summary = df_test %>%
  group_by(Variable) %>%
  summarize(
    Mean = mean(Value),
    SD = sd(Value),
    SE = SD / sqrt(n()),  # Standard Error
    Min = min(Value),
    Max = max(Value)
  )

# ggplot(df_test,aes(x=Variable,y=Value,group=Sample)) +
#   geom_line() + labs(
#     x = "Fragment Length (bp)",
#     y = "Proportion")


d = ggplot(df_test, aes(x=Variable, y=Value, group=Sample)) +
  geom_line(alpha=0.4) + 
  geom_line(data = df_summary, aes(x=Variable, y=Mean, group=1), size=1, color="black") + 
  labs(x = "Fragment Length (bp)", y = "Proportion")

ggarrange(a,b,c,d,ncol = 4,nrow=1)
```

## N- Category

```{r ,fig.width=15,fig.height=4}
#################
# N Category = N0 
#################
stat_tcat = df_OPC %>% 
  filter(stage_n_ajcc8 == "N0" & Timepoint == "Baseline") 

df_test = df_prop_jinfeng %>% 
  filter(Sample %in% stat_tcat$Sample) %>% 
  pivot_longer(cols = 2:202, names_to = "Variable", values_to = "Value") %>% 
  mutate(Variable = as.integer(Variable)) 

# Calculating means, standard deviations, and min/max values
df_summary = df_test %>%
  group_by(Variable) %>%
  summarize(
    Mean = mean(Value),
    SD = sd(Value),
    SE = SD / sqrt(n()),  # Standard Error
    Min = min(Value),
    Max = max(Value)
  )

# ggplot(df_test,aes(x=Variable,y=Value,group=Sample)) +
#   geom_line() + labs(
#     x = "Fragment Length (bp)",
#     y = "Proportion")


a = ggplot(df_test, aes(x=Variable, y=Value, group=Sample)) +
  geom_line(alpha=0.4) + 
  geom_line(data = df_summary, aes(x=Variable, y=Mean, group=1), size=1, color="black") + 
  labs(x = "Fragment Length (bp)", y = "Proportion")

#################
# N Category = N1 
#################
stat_tcat = df_OPC %>% 
  filter(stage_n_ajcc8 == "N1" & Timepoint == "Baseline") 

df_test = df_prop_jinfeng %>% 
  filter(Sample %in% stat_tcat$Sample) %>% 
  pivot_longer(cols = 2:202, names_to = "Variable", values_to = "Value") %>% 
  mutate(Variable = as.integer(Variable)) 

# Calculating means, standard deviations, and min/max values
df_summary = df_test %>%
  group_by(Variable) %>%
  summarize(
    Mean = mean(Value),
    SD = sd(Value),
    SE = SD / sqrt(n()),  # Standard Error
    Min = min(Value),
    Max = max(Value)
  )

# ggplot(df_test,aes(x=Variable,y=Value,group=Sample)) +
#   geom_line() + labs(
#     x = "Fragment Length (bp)",
#     y = "Proportion")


b = ggplot(df_test, aes(x=Variable, y=Value, group=Sample)) +
  geom_line(alpha=0.4) + 
  geom_line(data = df_summary, aes(x=Variable, y=Mean, group=1), size=1, color="black") + 
  labs(x = "Fragment Length (bp)", y = "Proportion")

#################
# N Category = N1 
#################
stat_tcat = df_OPC %>% 
  filter(stage_n_ajcc8 == "N2" & Timepoint == "Baseline") 

df_test = df_prop_jinfeng %>% 
  filter(Sample %in% stat_tcat$Sample) %>% 
  pivot_longer(cols = 2:202, names_to = "Variable", values_to = "Value") %>% 
  mutate(Variable = as.integer(Variable)) 

# Calculating means, standard deviations, and min/max values
df_summary = df_test %>%
  group_by(Variable) %>%
  summarize(
    Mean = mean(Value),
    SD = sd(Value),
    SE = SD / sqrt(n()),  # Standard Error
    Min = min(Value),
    Max = max(Value)
  )

# ggplot(df_test,aes(x=Variable,y=Value,group=Sample)) +
#   geom_line() + labs(
#     x = "Fragment Length (bp)",
#     y = "Proportion")


c = ggplot(df_test, aes(x=Variable, y=Value, group=Sample)) +
  geom_line(alpha=0.4) + 
  geom_line(data = df_summary, aes(x=Variable, y=Mean, group=1), size=1, color="black") + 
  labs(x = "Fragment Length (bp)", y = "Proportion")


#################
# N Category = N3 
#################
stat_tcat = df_OPC %>% 
  filter(stage_n_ajcc8 == "N3" & Timepoint == "Baseline") 

df_test = df_prop_jinfeng %>% 
  filter(Sample %in% stat_tcat$Sample) %>% 
  pivot_longer(cols = 2:202, names_to = "Variable", values_to = "Value") %>% 
  mutate(Variable = as.integer(Variable)) 

# Calculating means, standard deviations, and min/max values
df_summary = df_test %>%
  group_by(Variable) %>%
  summarize(
    Mean = mean(Value),
    SD = sd(Value),
    SE = SD / sqrt(n()),  # Standard Error
    Min = min(Value),
    Max = max(Value)
  )

# ggplot(df_test,aes(x=Variable,y=Value,group=Sample)) +
#   geom_line() + labs(
#     x = "Fragment Length (bp)",
    # y = "Proportion")


d = ggplot(df_test, aes(x=Variable, y=Value, group=Sample)) +
  geom_line(alpha=0.4) + 
  geom_line(data = df_summary, aes(x=Variable, y=Mean, group=1), size=1, color="black") + 
  labs(x = "Fragment Length (bp)", y = "Proportion")

ggarrange(a,b,c,d,ncol = 4,nrow=1)
```

## Overall Stage

```{r ,fig.width=15,fig.height=4}
#################
# Stage I 
#################
stat_tcat = df_OPC %>% 
  filter(stage_ajcc8 == "I" & Timepoint == "Baseline") 

df_test = df_prop_jinfeng %>% 
  filter(Sample %in% stat_tcat$Sample) %>% 
  pivot_longer(cols = 2:202, names_to = "Variable", values_to = "Value") %>% 
  mutate(Variable = as.integer(Variable)) 

# Calculating means, standard deviations, and min/max values
df_summary = df_test %>%
  group_by(Variable) %>%
  summarize(
    Mean = mean(Value),
    SD = sd(Value),
    SE = SD / sqrt(n()),  # Standard Error
    Min = min(Value),
    Max = max(Value)
  )

# ggplot(df_test,aes(x=Variable,y=Value,group=Sample)) +
#   geom_line() + labs(
#     x = "Fragment Length (bp)",
#     y = "Proportion")


a = ggplot(df_test, aes(x=Variable, y=Value, group=Sample)) +
  geom_line(alpha=0.4) + 
  geom_line(data = df_summary, aes(x=Variable, y=Mean, group=1), size=1, color="black") + 
  labs(x = "Fragment Length (bp)", y = "Proportion")

#################
# Stage II
#################
stat_tcat = df_OPC %>% 
  filter(stage_ajcc8 == "II" & Timepoint == "Baseline") 

df_test = df_prop_jinfeng %>% 
  filter(Sample %in% stat_tcat$Sample) %>% 
  pivot_longer(cols = 2:202, names_to = "Variable", values_to = "Value") %>% 
  mutate(Variable = as.integer(Variable)) 

# Calculating means, standard deviations, and min/max values
df_summary = df_test %>%
  group_by(Variable) %>%
  summarize(
    Mean = mean(Value),
    SD = sd(Value),
    SE = SD / sqrt(n()),  # Standard Error
    Min = min(Value),
    Max = max(Value)
  )

# ggplot(df_test,aes(x=Variable,y=Value,group=Sample)) +
#   geom_line() + labs(
#     x = "Fragment Length (bp)",
#     y = "Proportion")


b = ggplot(df_test, aes(x=Variable, y=Value, group=Sample)) +
  geom_line(alpha=0.4) + 
  geom_line(data = df_summary, aes(x=Variable, y=Mean, group=1), size=1, color="black") + 
  labs(x = "Fragment Length (bp)", y = "Proportion")

#################
# Stage III
#################
stat_tcat = df_OPC %>% 
  filter(stage_ajcc8 == "III" & Timepoint == "Baseline") 

df_test = df_prop_jinfeng %>% 
  filter(Sample %in% stat_tcat$Sample) %>% 
  pivot_longer(cols = 2:202, names_to = "Variable", values_to = "Value") %>% 
  mutate(Variable = as.integer(Variable)) 

# Calculating means, standard deviations, and min/max values
df_summary = df_test %>%
  group_by(Variable) %>%
  summarize(
    Mean = mean(Value),
    SD = sd(Value),
    SE = SD / sqrt(n()),  # Standard Error
    Min = min(Value),
    Max = max(Value)
  )

# ggplot(df_test,aes(x=Variable,y=Value,group=Sample)) +
#   geom_line() + labs(
#     x = "Fragment Length (bp)",
#     y = "Proportion")


c = ggplot(df_test, aes(x=Variable, y=Value, group=Sample)) +
  geom_line(alpha=0.4) + 
  geom_line(data = df_summary, aes(x=Variable, y=Mean, group=1), size=1, color="black") + 
  labs(x = "Fragment Length (bp)", y = "Proportion")

ggarrange(a,b,c,ncol=3)
```

# Fragment Length

## Fragment length proportion by group.

```{r}
df_under150a <- df_prop_jinfeng %>%
  # Calculate the sum of columns 3 through 603
  mutate(hpv_prop_sum = rowSums(.[, 3:102])) %>% 
  select(Sample,hpv_prop_sum) %>% 
  left_join(select(df_OPC,Sample,patient,Timepoint,stage_n_ajcc8,stage_t_ajcc8,stage_ajcc8),by = "Sample") %>% 
  filter(Timepoint == "Baseline" & !is.na(stage_n_ajcc8))

df_under150 <- df_prop_healthy %>%
  mutate(hpv_prop_sum = rowSums(.[, 3:102])) %>% 
  select(Sample,hpv_prop_sum) %>% 
  left_join(select(df_OPC,Sample,patient,Timepoint,stage_n_ajcc8,stage_t_ajcc8,stage_ajcc8),by = "Sample") %>% 
  mutate(stage_n_ajcc8 = "HUCON",
         stage_t_ajcc8 = "HUCON",
         stage_ajcc8= "HUCON") %>% 
  rbind(df_under150a)

ggplot(df_under150, aes(x=stage_n_ajcc8, y=hpv_prop_sum)) +
  geom_boxplot() + stat_compare_means() + geom_jitter(alpha=0.3)

ggplot(df_under150, aes(x=stage_t_ajcc8, y=hpv_prop_sum)) +
  geom_boxplot() + stat_compare_means() + geom_jitter(alpha=0.3)


ggplot(df_under150, aes(x=stage_ajcc8, y=hpv_prop_sum)) +
  geom_boxplot() + stat_compare_means() + geom_jitter(alpha=0.3)
```

# Fragment length proportion over time

```{r}

df_under150 <- df_prop_jinfeng %>%
  # Calculate the sum of columns 3 through 603
  mutate(hpv_prop_sum = rowSums(.[, 3:102])) %>% 
  select(Sample,hpv_prop_sum) %>% 
  left_join(select(df_OPC,Sample,patient,Timepoint,seq_copies_mL_corrected1.67,death,progression),by = "Sample")

# fwrite(df_under150,"~/Desktop/human_fragprop_under150bp.txt",quote=F,row.names = F,sep="\t",col.names = T)

# Filter out patients without a baseline measurement first
filtered_df <- df_under150 %>%
  group_by(patient) %>%
  filter("Baseline" %in% Timepoint)

# Then, for the remaining patients, compute the ratios
result <- filtered_df %>%
  mutate(baseline_hpv = hpv_prop_sum[Timepoint == "Baseline"]) %>%
  mutate(ratio = hpv_prop_sum / baseline_hpv) %>%
  select(-baseline_hpv) %>%  # Optionally remove the baseline column
  mutate(Timepoint = case_when(
    Timepoint == "Baseline" ~ 1,
    Timepoint == "Mid RT" ~ 2,
    Timepoint == "Early RT" ~ 3,
    Timepoint == "3 Month" ~ 4,
    TRUE ~ 5
  ))

stat_tcat = df_OPC %>%
  filter(progression == TRUE)

df_test = result %>% 
  filter(patient %in% stat_tcat$patient) %>% 
  left_join(select(df_OPC,death,stage_n_ajcc8,Sample),by="Sample")

ggplot(df_test, aes(x=Timepoint, y=ratio, group=patient)) +
  geom_line(alpha=0.5) +
  labs(x = "Time", y = "Proportion change since baseline")+
  facet_wrap(~stage_n_ajcc8)

# ggplot(df_test, aes(x=Timepoint, y=hpv_prop_sum, group=patient)) +
#   geom_line(alpha=0.5) +
#   labs(x = "Time", y = "Proportion <150bp")+
#   facet_wrap(~stage_n_ajcc8)

ggplot(df_test, aes(x=Timepoint, y=seq_copies_mL_corrected1.67, group=patient)) +
  scale_y_log10() +
  geom_line(alpha=0.5) +
  labs(x = "Time", y = "ctDNA Concentration") +
  facet_wrap(~stage_n_ajcc8)
```

# Association with GTV and Fragment proportion

```{r, warning = FALSE, fig.width= 8, fig.height = 3}
df_under150_healthy = df_prop_healthy %>% 
  # Calculate the sum of columns 3 through 603
  mutate(hpv_prop_sum = rowSums(.[, 3:102])) %>% 
  select(Sample,hpv_prop_sum) %>% 
  mutate(gtv_combined = 0, 
         sum_gtv_nodes = 0,
         gtv_primary = 0,
         seq_copies_mL_corrected1.67 = 0) %>% 
  mutate(Group = "Healthy") %>% 
  select(-Sample)

df_under150 <- df_prop_jinfeng %>%
  mutate(hpv_prop_sum = rowSums(.[, 3:102])) %>% 
  select(Sample,hpv_prop_sum) %>% 
  left_join(select(df_OPC,Sample,patient,Timepoint,seq_copies_mL_corrected1.67,sum_gtv_nodes,gtv_primary),by = "Sample") %>% 
  filter(Timepoint =="Baseline") %>% 
  mutate(gtv_combined = sum_gtv_nodes + gtv_primary) %>% 
  select(hpv_prop_sum,seq_copies_mL_corrected1.67,sum_gtv_nodes,gtv_primary,gtv_combined) %>% 
  mutate(gtv_combined = log1p(gtv_combined),
         sum_gtv_nodes = log1p(sum_gtv_nodes),
         gtv_primary = log1p(gtv_primary),
         seq_copies_mL_corrected1.67 = log1p(seq_copies_mL_corrected1.67)) %>%
  mutate(Group = "Baseline") %>% 
  rbind(df_under150_healthy)



p_gtv1 = ggplot(df_under150, aes(x=gtv_primary, y=hpv_prop_sum,color=Group)) +
  geom_point() + 
  labs(x = "log(GTV primary)", y = "Proportion of Fragments <150bp")

p_gtv2 = ggplot(df_under150, aes(x=sum_gtv_nodes, y=hpv_prop_sum,color=Group)) +
  geom_point() + 
  labs(x = "log(GTV nodes)", y = "Proportion of Fragments <150bp")

p_gtv3 = ggplot(df_under150, aes(x=gtv_combined, y=hpv_prop_sum,color=Group)) +
  geom_point() + 
  labs(x = "log(GTV combined)", y = "Proportion of Fragments <150bp")

ggarrange(p_gtv1,p_gtv2,p_gtv3,ncol=3,common.legend = TRUE)

# ggpairs(df_under150,progress = FALSE)
```

# Simple linear model with Fragment length and ctDNA concentration

```{r}
# df_under150
# sum_gtv_nodes+  gtv_primary
lm_ctdna = lm( gtv_combined~ hpv_prop_sum,data = df_under150)
summary(lm_ctdna)

df_subset_prop_gtv <- df_under150 %>% 
  select(-Group) 

ggpairs(df_subset_prop_gtv)
  
```

