---
title: "Fragmentomic Analysis"
output: html_document
date: "2023-11-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(caret)
library(DataExplorer)
library(data.table)
library(dplyr)
library(faraway)
library(forestmodel)
library(ggbeeswarm)
library(ggplot2)
library(ggpubr)
# library(ggthemr)
library(gridExtra)
library(NMF)
library(rlang)
library(survival)
library(survminer)
library(readxl)
library(tidyverse)
library(viridis)



```

```{r}
###########
# Read OPC
###########

df_OPC <- read_excel("~/Documents/1 Bratman/head_neck_analysis_lucas/data/Quantification/summ_OPC.v6.xlsx")
colnames(df_OPC)[1] = c("filename")
colnames(df_OPC)[19] = c("timepoint")
df_OPC$timepoint <- gsub("-", "", df_OPC$timepoint)
# 
# reseq_remove = readLines("/Users/lucaspenny/Documents/1 Bratman/head_neck_analysis_lucas/data/re-sequenced-remove.txt")
# 
# reseq_remove = df_OPC %>%
#   filter(`Library Name` %in% reseq_remove) %>%
#   select(`Library Name`,filename)


################
# read frags in and filter 
################

df_hg_frag = fread("/Users/lucaspenny/Documents/1 Bratman/head_neck_analysis_lucas/data/20231115_hg19_fragmentSize_LucasUpdated_pt1.csv")
df_hg_frag1 = fread("/Users/lucaspenny/Documents/1 Bratman/head_neck_analysis_lucas/data/20231115_hg19_fragmentSize_LucasUpdated_pt2.csv")
df_hg_frag_healthy = fread("/Users/lucaspenny/Documents/1 Bratman/head_neck_analysis_lucas/data/fragment_length_summary_healthy.csv")
df_virus_frag = fread("/Users/lucaspenny/Documents/1 Bratman/head_neck_analysis_lucas/data/20231115_virus_fragmentSize_LucasUpdated_pt1.csv") 
df_virus_frag1 = fread("/Users/lucaspenny/Documents/1 Bratman/head_neck_analysis_lucas/data/20231120_virus_reseq_fragment_length_summary.csv") 

df_hg_frag = df_hg_frag %>% 
  rbind(df_hg_frag1)

df_OPC <- bind_rows(df_OPC, select(df_hg_frag_healthy, filename))



########
# split for virus 
########
# Function to split and extract the desired parts
split_and_extract <- function(x) {
  split_1 <- unlist(strsplit(x, "_hgUnmapped_"))
  part_1 <- split_1[1]
  part_2 <- unlist(strsplit(split_1[2], "\\."))[1]
  return(c(part_1, part_2))
}

# Apply the function to each element of the column and create two new columns
df_virus_frag$virus <- sapply(df_virus_frag$filename, function(x) split_and_extract(x)[2])
df_virus_frag$filename <- sapply(df_virus_frag$filename, function(x) split_and_extract(x)[1])


#########
# split for human
########

df_hg_frag$filename <- gsub("_output_met.*", "", df_hg_frag$filename)
df_virus_frag1$filename <- gsub("_metrics.txt", "", df_virus_frag1$filename)
df_virus_frag1$virus = "HPV16Aligned"

df_virus_frag = df_virus_frag %>% 
  rbind(df_virus_frag1)

########
# filter out by bad sequencing 

df_hg_frag = df_hg_frag %>% 
  filter(filename %in% df_OPC$filename) %>% 
  rbind(df_hg_frag_healthy)

df_virus_frag = df_virus_frag %>% 
  filter(filename %in% df_OPC$filename)
                     
                     


##########
# read clinical
##########

df_clinical = fread("/Users/lucaspenny/Documents/1 Bratman/head_neck_analysis_lucas/data/clinical_data.tsv")
df_clinical = df_clinical %>%
  mutate(Patient_ID = patient) %>% 
  left_join(df_OPC %>% dplyr::select(filename,Patient_ID))


attach('/Users/lucaspenny/Documents/1 Bratman/head_neck_analysis_lucas/HPV_analysis_dataframes.Rdata'); df_gtv_nodes <- df_gtv_nodes
attach('/Users/lucaspenny/Documents/1 Bratman/head_neck_analysis_lucas/HPV_analysis_dataframes.Rdata'); df_gtv_primary <- df_gtv_primary

df_clinical = df_clinical %>%
  mutate(Patient_ID = patient) %>% 
  left_join(df_OPC %>% dplyr::select(filename,Patient_ID)) %>% 
  left_join(df_gtv_nodes, by = "patient") %>% 
  left_join(df_gtv_primary, by = "patient")
```

```{r}

###########
# Process the files
###########

df_prop_virus <- df_virus_frag %>%
  # Calculate the sum of columns 3 through 603
  mutate(total_sum = rowSums(.[, 3:603])) %>%
  filter(virus == "HPV16Aligned") %>% 
  dplyr::select(filename, everything()) %>%
  left_join(dplyr::select(df_OPC, filename, timepoint), by = "filename") %>% 
  filter(timepoint == "Baseline"  & READ_PAIRS > 500) %>%  #POSSIBLY CHANGE
  # Divide each column by the respective total_sum
  mutate(across(3:603, ~ . / total_sum)) %>% 
  dplyr::select(-READ_PAIRS, -V604, -timepoint,-total_sum,-virus) %>%
  arrange(filename)

df_prop_hg <- df_hg_frag %>%
  # Calculate the sum of columns 3 through 603
  mutate(total_sum = rowSums(.[, 3:603])) %>%
  dplyr::select(filename, everything()) %>%
  left_join(dplyr::select(df_OPC, filename, timepoint), by = "filename") %>% 
  filter(timepoint == "Baseline"| is.na(timepoint)) %>% 
  # Divide each column by the respective total_sum
  mutate(across(3:603, ~ . / total_sum)) %>% 
  dplyr::select(-READ_PAIRS, -V604, -timepoint,-total_sum) %>%
  # filter(filename %in% df_prop_virus$filename) %>%
  arrange(filename)

# df_prop_subtracted <- df_prop_hg %>% 
#   # substract column 2:602 from df_prop_virus
#   mutate(across(2:602, ~ . - df_prop_virus[[cur_column()]]))
df_prop_subtracted <- df_prop_hg %>% 
  mutate(across(2:602, ~ abs(. - df_prop_virus[[cur_column()]])))

##############
# find longest frag per row
##############
find_last_nonzero_column <- function(row) {
  # Get the index of the last non-zero value
  last_nonzero_index <- max(which(row > 0.0000001))
  
  # If there are no non-zero values, return NA
  if (length(last_nonzero_index) == 0) {
    return(NA)
  } else {
    # Return the column name
    return(names(row)[last_nonzero_index])
  }
}

# Apply the function to each row
result <- apply(df_prop_hg[, 50:600], 1, find_last_nonzero_column)
hist(as.numeric(result))
shortest_common_max <- names(sort(table(as.numeric(result)), decreasing = TRUE))[1] # 265 for virus, 364 human

result <- apply(df_prop_virus[, 50:600], 1, find_last_nonzero_column)
hist(as.numeric(result))
shortest_common_max <- names(sort(table(as.numeric(result)), decreasing = TRUE))[1] # 265 for virus, 364 human

################
# re-calculate the proportion accordingly 
################

df_prop_virus <- df_virus_frag %>%
  # Calculate the sum of columns 3 through 603
  mutate(total_sum = rowSums(.[, 53:268])) %>%
  filter(virus == "HPV16Aligned") %>% 
  dplyr::select(filename, everything()) %>%
  left_join(dplyr::select(df_OPC, filename, timepoint), by = "filename") %>% 
  filter((timepoint == "Baseline" | is.na(timepoint)) & READ_PAIRS > 1000) %>%  #POSSIBLY CHANGE
  # Divide each column by the respective total_sum
  mutate(across(53:268, ~ . / total_sum)) %>% 
  dplyr::select(filename,53:268) %>%
  arrange(filename)

df_prop_hg <- df_hg_frag %>%
  mutate(total_sum = rowSums(.[, 53:367])) %>%
  dplyr::select(filename, everything()) %>%
  left_join(dplyr::select(df_OPC, filename, timepoint), by = "filename") %>% 
  filter((timepoint == "Baseline" | is.na(timepoint)) & READ_PAIRS > 1000) %>% 
  # Divide each column by the respective total_sum
  mutate(across(53:367, ~ . / total_sum)) %>% 
  dplyr::select(filename,53:367) %>%
  # filter(filename %in% df_prop_virus$filename) %>%
  arrange(filename)




```

## Visualize Fragment length distributions

```{r}
ggthemr::ggthemr_reset()


##################
# virus reads
##################
maxmima_vals = df_prop_virus[order(df_prop_virus$`175`), ][1, ]$filename


df_fraglength_prop_long <- df_prop_virus %>%
  # filter(str_detect(filename, "OPC_Pool-1_1_S1_L001")) %>%
  pivot_longer(cols = -c(filename), names_to = "length", values_to = "proportion") %>%
  mutate(length = as.numeric(length))


#plot all sample fragment length proportions
ggplot(df_fraglength_prop_long, aes(x = length, y = proportion, group = filename, color = filename)) + 
  geom_line(alpha = 0.7) +
  labs(x = "Length", y = "Proportion") +
  theme_minimal() + theme(legend.position = "none") +
  xlim(0,300) + ggtitle("Fragment Length Distribution of Virus-Only Reads")

##################
# summary stats

df_summary <- df_fraglength_prop_long %>%
  group_by(length) %>%
  summarise(
    # median_proportion = median(proportion),
    mean_proportion = mean(proportion),
    sd_proportion = sd(proportion)
  )

ggplot(df_summary, aes(x = length, y = mean_proportion)) +
  geom_line() +
  geom_ribbon(aes(ymin = mean_proportion - sd_proportion, ymax = mean_proportion + sd_proportion), alpha = 0.2) +
  labs(x = "Length", y = "Proportion", title = "Proportion by Length with Standard Deviation") +
  theme_minimal()


##################
# human reads
##################
outlier_x = df_prop_hg[which.max(df_prop_hg$`130`),1] 
minima_vals = df_prop_hg[order(df_prop_hg$`175`), ][1:3, ]$filename

summary_minima = df_clinical %>% 
  # filter(filename %in% minima_vals) %>% 
  summarize(stage_t_ajcc8 = stage_t_ajcc8,
            stage_n_ajcc8 = stage_n_ajcc8,
            stage_ajcc8 = stage_ajcc8,
            sum_gtv_nodes.x = sum_gtv_nodes,
            gtv_primary = gtv_primary)


df_fraglength_prop_long <- df_prop_hg %>% 
  filter(filename ) %>%
  pivot_longer(cols = -c(filename), names_to = "length", values_to = "proportion") %>%
  mutate(length = as.numeric(length))

#plot all sample fragment length proportions
ggplot(df_fraglength_prop_long, aes(x = length, y = proportion, group = filename, color = filename)) + 
  geom_line(alpha = 0.7) +
  labs(x = "Length", y = "Proportion") +
  theme_minimal() + theme(legend.position = "none") + 
  xlim(0,300) + ggtitle("Fragment Length Distribution of Human-Only Reads")


##################
# summary stats

df_summary <- df_fraglength_prop_long %>%
  group_by(length) %>%
  summarise(
    # median_proportion = median(proportion),
    mean_proportion = mean(proportion),
    sd_proportion = sd(proportion)
  )

ggplot(df_summary, aes(x = length, y = mean_proportion)) +
  geom_line() +
  geom_ribbon(aes(ymin = mean_proportion - sd_proportion, ymax = mean_proportion + sd_proportion), alpha = 0.2) +
  labs(x = "Length", y = "Proportion", title = "Proportion by Length with Standard Deviation") +
  theme_minimal()



##################
# human reads (HEALTHY ONLY)
##################


df_fraglength_prop_long <- df_prop_hg %>% 
  filter(str_detect(df_prop_hg$filename, "HUCON")) %>%
  pivot_longer(cols = -c(filename), names_to = "length", values_to = "proportion") %>%
  mutate(length = as.numeric(length))

#plot all sample fragment length proportions
ggplot(df_fraglength_prop_long, aes(x = length, y = proportion, group = filename, color = filename)) + 
  geom_line(alpha = 0.7) +
  labs(x = "Length", y = "Proportion") +
  theme_minimal() + theme(legend.position = "none") + 
  xlim(0,300) + ggtitle("Fragment Length Distribution of Human-Only Reads")


##################
# summary stats

df_summary <- df_fraglength_prop_long %>%
  group_by(length) %>%
  summarise(
    # median_proportion = median(proportion),
    mean_proportion = mean(proportion),
    sd_proportion = sd(proportion)
  )

ggplot(df_summary, aes(x = length, y = mean_proportion)) +
  geom_line() +
  geom_ribbon(aes(ymin = mean_proportion - sd_proportion, ymax = mean_proportion + sd_proportion), alpha = 0.2) +
  labs(x = "Length", y = "Proportion", title = "Proportion by Length with Standard Deviation") +
  theme_minimal()
```

### Further virus analysis

```{r}
##################
# virus reads
##################
maxmima_vals = df_prop_virus[order(df_prop_virus$`175`), ][1, ]$filename

x = (df_OPC %>% dplyr::filter(IntegrationLow>10) %>% dplyr::select(filename)) #change to what filter I want
df_fraglength_prop_long <- df_prop_virus %>%
  filter(filename %in% x$filename) %>%
  pivot_longer(cols = -c(filename), names_to = "length", values_to = "proportion") %>%
  mutate(length = as.numeric(length))


#plot all sample fragment length proportions
ggplot(df_fraglength_prop_long, aes(x = length, y = proportion, group = filename, color = filename)) + 
  geom_line(alpha = 0.7) +
  labs(x = "Length", y = "Proportion") +
  theme_minimal() + theme(legend.position = "none") +
  xlim(0,300) + ggtitle("Fragment Length Distribution of Virus-Only Reads")

##################
# summary stats

df_summary <- df_fraglength_prop_long %>%
  group_by(length) %>%
  summarise(
    # median_proportion = median(proportion),
    mean_proportion = mean(proportion),
    sd_proportion = sd(proportion)
  )

ggplot(df_summary, aes(x = length, y = mean_proportion)) +
  geom_line() +
  geom_ribbon(aes(ymin = mean_proportion - sd_proportion, ymax = mean_proportion + sd_proportion), alpha = 0.2) +
  labs(x = "Length", y = "Proportion", title = "N3 Proportion by Length with Standard Deviation") +
  theme_minimal()
```

## Create NMF Graphs

```{r}

df_clinical = fread("/Users/lucaspenny/Documents/1 Bratman/head_neck_analysis_lucas/data/clinical_data.tsv")

df_clinical = df_clinical %>% 
  mutate(Patient_ID = patient)

###############
# Create Matrix for NMF
###############
m_all_insertsize_extended =t(as.matrix(df_prop_virus[, -1])) #df_prop_hg
colnames(m_all_insertsize_extended) = (df_prop_virus$filename)

# m_all_insertsize_extended =t(as.matrix(df_prop_hg[, -1])) #df_prop_hg
# colnames(m_all_insertsize_extended) = (df_prop_hg$filename)

###############
# Create df for NMF analysis 
###############

df_samples <- df_OPC %>%
  left_join(df_clinical,by = "Patient_ID") %>% 
  dplyr::select(library = filename, 
                sample = stage_n_ajcc8, 
                patient = Patient_ID) %>%
  # filter(sample == "N2" | str_detect(library, "HUCON")) %>%
  mutate(cohort = "OPC",
         timepoint = "Baseline")

nmf_insertsize <- nmf(
  m_all_insertsize_extended,
  rank = 2
)

plotdata_nmf_basis <- basis(nmf_insertsize) %>%
  as_tibble(rownames = 'fragment_length') %>%
  gather(component, basis, -fragment_length) %>%
  mutate(fragment_length = as.integer(fragment_length)) %>%
  group_by(component) %>%
  mutate(
    proportion = basis / sum(basis),
    component_mean = sum(fragment_length * proportion)
  ) %>%
  ungroup() %>%
  mutate(
    signature = if_else(component_mean == min(component_mean), 'cancer', 'normal')
  ) %>%
  ungroup() %>%
  dplyr::select(-component_mean)

plotdata_nmf_weights <- coef(nmf_insertsize) %>%
  t() %>%
  as_tibble(rownames = 'library') %>%
  gather(component, value, -library) %>%
  left_join(plotdata_nmf_basis %>% distinct(component, signature)) %>%
  dplyr::select(-component) %>%
  spread(signature, value) %>%
  right_join(df_samples) %>% #change back to left join 
  mutate(
  sample_type = sample,
  sample_type = factor(sample_type, levels = c("N0", "N1", "N2", "N3", NA))) %>%
  # sample_type = factor(sample_type, levels = c("T1", "T2", "T3","T4", NA))) %>%
  dplyr::select(library, sample_type, cancer, normal) %>%
  arrange(sample_type, cancer) %>%
  mutate(
    library = factor(
      library,
      levels = unique(library)
    )
  ) %>%
  gather(component, weight, -library, -sample_type) %>%
  group_by(library) %>%
  mutate(weight = weight / sum(weight)) %>%
  ungroup()

p_heatmap_nmf_weights <- plotdata_nmf_weights %>%
  ggplot(aes(
    x = component,
    y = library,
    fill = weight
  )) +
  geom_tile() +
  scale_fill_viridis() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )

p_colorbar_nmf_weights <- plotdata_nmf_weights %>%
  ggplot(aes(
    x = 0,
    y = library,
    fill = sample_type
  )) +
  geom_tile() +
  scale_fill_brewer(palette = 'Set1') +
  theme_void()

p_nmf_weights_beeswarm <- plotdata_nmf_weights %>%
  filter(component == 'cancer') %>%
  ggplot(aes(
    x = sample_type,
    y = weight
  )) +
  geom_beeswarm(color = 'black', size = 0.5) +
  labs(
    x = 'Sample Type',
    y = 'NMF Cancer\nComponent Weight'
  )

# plotdata_nmf_immunescore <- plotdata_nmf_weights %>%
#   inner_join(
#     bind_rows(
#       df_csm,
#       df_csm_normalcontrol
#     ),
#     by = 'library'
#   )
# 
# p_nmf_csm_correlation <- plotdata_nmf_immunescore %>%
#   filter(immune_group == 'Non-immune') %>%
#   ggplot(aes(
#     x = weight,
#     y = score,
#     color = component
#   )) +
#   facet_grid(component ~ .) +
#   geom_point() +
#   scale_y_log10() +
#   labs(
#     x = 'NMF component weight',
#     y = 'CSM'
#   )

plotdata_nmf_cancer_likelihood <- plotdata_nmf_basis %>%
  left_join(
    plotdata_nmf_basis %>%
      group_by(component) %>%
      summarise(mean_fl = sum(fragment_length * proportion)) %>%
      mutate(component_name = if_else(mean_fl == min(mean_fl), 'Cancer', 'Normal'))
  ) %>%
  dplyr::select(component_name, fragment_length, proportion) %>%
  spread(component_name, proportion) %>%
  mutate(cancer_likelihood = Cancer / (Cancer + Normal))

p_nmf_basis <- plotdata_nmf_cancer_likelihood %>%
  gather(group, value, -fragment_length, -cancer_likelihood) %>%
  ggplot(aes(
    x = fragment_length,
    y = value,
    color = group,
    group = group
  )) +
  geom_line() +
  labs(
    x = 'Fragment Length',
    y = 'Basis value',
    color = 'NMF component'
  )

p_nmf_basis_likelihood <- plotdata_nmf_cancer_likelihood %>%
  distinct(fragment_length, cancer_likelihood) %>%
  ggplot(aes(
    x = fragment_length,
    y = cancer_likelihood
  )) +
  geom_line()


plot(p_heatmap_nmf_weights)

plot(p_colorbar_nmf_weights)
plot(p_nmf_weights_beeswarm)
plot(p_nmf_basis)
plot(p_nmf_basis_likelihood)


```

-   take all the frags in a sample and map to the cancer likelihood, and then you can sum that up and see how cancer

### Curious about swapping the cancer and non-cancer info

```{r}
plotdata_nmf_cancer_likelihood <- plotdata_nmf_basis %>%
  left_join(
    plotdata_nmf_basis %>%
      group_by(component) %>%
      summarise(mean_fl = sum(fragment_length * proportion)) %>%
      mutate(component_name = if_else(mean_fl == min(mean_fl), 'Cancer', 'Normal'))
  ) %>%
  dplyr::select(component_name, fragment_length, proportion) %>%
  spread(component_name, proportion) %>%
  mutate(cancer_likelihood = Normal / (Cancer + Normal))

p_nmf_basis <- plotdata_nmf_cancer_likelihood %>%
  gather(group, value, -fragment_length, -cancer_likelihood) %>%
  ggplot(aes(
    x = fragment_length,
    y = value,
    color = group,
    group = group
  )) +
  geom_line() +
  labs(
    x = 'Fragment Length',
    y = 'Basis value',
    color = 'NMF component'
  )

p_nmf_basis_likelihood <- plotdata_nmf_cancer_likelihood %>%
  distinct(fragment_length, cancer_likelihood) %>%
  ggplot(aes(
    x = fragment_length,
    y = cancer_likelihood
  )) +
  geom_line()


plot(p_heatmap_nmf_weights)

plot(p_colorbar_nmf_weights)
plot(p_nmf_weights_beeswarm)
plot(p_nmf_basis)
plot(p_nmf_basis_likelihood)
```

## Compare the fragment score that we talked about in lab meeting

Althaf mentioned multiplying the likelihood by the given patient matrix.

-   I tried for both virus and human ( just need to change the line for m_all_insert_size variable above and run to below to get the result. It doesn't appear that there are any significant differences of note here for human. Very unfortunate.
-   Can maybe try for the healthy vs normal signature though next.

```{r}

df_likelihood_filter = plotdata_nmf_cancer_likelihood %>% 
  filter(cancer_likelihood >0.66)

df_fraglength_prop_long <- df_prop_virus %>% pivot_longer(cols = -c(filename), 
                                                       names_to = "fragment_length", 
                                                       values_to = "proportion") %>%
  mutate(fragment_length = as.integer(fragment_length)) %>% 
  left_join(df_likelihood_filter,by = c("fragment_length"))


df_fraglength_prop_long$product <- df_fraglength_prop_long$proportion * df_fraglength_prop_long$cancer_likelihood

df_summary_likelihood <- df_fraglength_prop_long %>%
  group_by(filename) %>%
  summarise(total_product = sum(product, na.rm = TRUE)) %>% 
  right_join(df_OPC, by = "filename") %>% 
  left_join(df_clinical, by =c("Patient_ID")) %>% 
  left_join(df_gtv_primary,by="patient") %>% 
  left_join(df_gtv_nodes,by="patient") 


x_var = df_summary_likelihood$total_product
y_var = df_summary_likelihood$sum_gtv_nodes #gtv_primary #
# x_var = log(df_summary_likelihood$seq_copies_mL_corrected1.67+1)
p_scatter_weights_integration <- df_summary_likelihood %>%
  ggplot(aes(
    x = x_var, #log(seq_copies_mL_corrected1.67+1)
    y = y_var,
  )) +
  geom_point() # +  scale_fill_viridis() 

plot(p_scatter_weights_integration)


lm_seq_product = lm(y_var ~ x_var, data=df_summary_likelihood)

summary(lm_seq_product)

cor.test(x_var, y_var, method="pearson")
```

## Check Fragment length (90-150bp) prop and association

```{r}
# Apply the function to each row
# result <- apply(df_prop_virus[, 50:600], 1, find_last_nonzero_column)
# hist(as.numeric(result))
# shortest_common_max <- names(sort(table(as.numeric(result)), decreasing = TRUE))[1] # 265 for virus, 364 human

```

## Compare NMF weight with integration confidence

You could also use the NMF weights matrix to see whether cases with a higher weight from the shorter viral fragment length profile are also more likely to have high confidence integration events

-   plot out the fragment length profiles instead of just filtering out right away

```{r}

df_weights_OPC <- df_OPC %>%
  left_join(plotdata_nmf_weights, by = c("filename" = "library")) %>%
  filter(component == "cancer" & weight > 0.7)

df_weights_clinical <- df_clinical %>%
  left_join(df_weights_OPC, by = c("Patient_ID")) %>%
  filter(component == "cancer" & weight > 0.7) %>% 
  left_join(df_gtv_nodes, by = c("patient"))

df_fragpropvirus_integration <- df_prop_virus %>% 
  mutate(total_sum = rowSums(.[, 160:170]),
         total_sum2 = rowSums(.[, 130:140]),
         propval = log(total_sum)/log(total_sum2)) %>%
  select(filename,propval) %>% 
  left_join(df_OPC %>% dplyr::select(filename,Integration_score)) %>% 
  left_join(df_clinical %>% dplyr::select(filename,stage_n_ajcc8)) 
  
p_scatter_weights_integration <- df_weights_clinical %>%
  ggplot(aes(
    x = weight,
    y = sum_gtv_nodes
  )) +
  geom_point() +  scale_fill_viridis() 

p_scatter_weights_integration <- df_fragpropvirus_integration %>%
  ggplot(aes(
    x = propval,
    y = Integration_score
  )) +
  geom_point() +  scale_fill_viridis() 
p_scatter_weights_integration

plot(p_scatter_weights_integration)
```

## Random

```{r}

```
